---
title: "Technical Appendix Midterm"
author: "Mark Russeff & Loi Pham"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---

```{r}
# Course: BUAN 5210
# Title: Technical Appendix Midterm
# Purpose: Midterm Project 
# Date: February 14, 2019
# Authors: Mark Russeff & Loi Pham
```


```{r echo = FALSE, message=FALSE, warning=FALSE}

# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 

# Clear environmet of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE)lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), detach, character.only = TRUE, unload = TRUE)

```
```{r message = FALSE, warning = FALSE}

# Load packages
library(tidyverse)
library(GGally)
library(gridExtra)
library(kableExtra)
library(scales)
library(cowplot)
```

#Load and Review Data
```{r}
#Load data
mtp_product <- read.csv("mtp_product_data.csv")
mtp_sales <- read.csv("mtp_sales_data.csv")

#Check structure of the data
str(mtp_product)
str(mtp_sales)
```

##Join the Data Sets
```{r}
#Modify strings in UPC for join
mtp_product$UPC <- str_replace(mtp_product$UPC,"00-","")
mtp_sales$UPC <- str_replace_all(mtp_sales$UPC,"[.]","-")
  
#Join data frames into one by UPC
mtp <- mtp_sales %>%
  full_join(mtp_product, by = "UPC")

```

##Clean and Organize the Joined Data
```{r}
companies <- c("GENERAL MILLS", "KELLOGGS", "POST")

mtp <- mtp %>% 
  mutate(company = brand, cereal = brand)

for (pattern in companies) {
  mtp <- mtp %>% 
    mutate(company = if_else(str_detect(company, pattern),
                             pattern, as.character(company)),
           cereal = str_replace(cereal, pattern, ""), 
           cereal = str_replace(cereal, "TST CR", "TOAST CRUNCH"))
}

mtp <- mtp %>% 
  mutate(company = as.factor(str_to_title(company)),
         company = str_trim(str_replace(company, " ", "_")),
         cereal = as.factor(str_trim(str_to_title(cereal))),
         promo = as.factor(promo),
         
         revenue = price * units,
         price_volume = price / volume) %>% 
  select(company, cereal, price, units, volume, revenue, price_volume, promo, ad, package, week, iri_key)
```
# Univariate non-graphical

##Categorical data - summary
```{r}
# Inspect descriptive statistics of data
summary(mtp)

```

**Observations on the data**

- Kelloggs sells the most, followed by General Mills and Post.
- Kelloggs Frosted Flakes and Froot Loops are the most popular cereals. 
- Cinnamon Toast Crunch is General Mills' most popular cerearl.
- Price has a tight interquartile range but a very low min and a high max.
- Revenue is skewed due to some very large numbers.
- Because it is calculated from price, price per volume also has a tight interquartile range and high min/max values.
- Roughly 20% of the time there is an in-store promo.
- Only about 11.5% of the time are there ads being run.
- Ads and Promos make up a small percentage of total sales transactions.


##Categorical data - tabulation

###Company
```{r}
#Table of data by company.
mtp %>% 
  group_by(company) %>% 
  summarise(count = n(),
  mean_revenue = mean(revenue),
  mean_price = mean(price),
  mean_price_vol = mean(price_volume),
  mean_units = mean(units)) %>%
  arrange(desc(mean_revenue)) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

* Kelloggs has the highest number of sales but General Mills has the highest average revenue.
* It looks like General Mills may have more pricing power in the market.  

###Cereal

```{r}
#Table of data by cereal
mtp %>% 
  group_by(company, cereal) %>% 
  summarise(count = n(),
  mean_revenue = mean(revenue),
  mean_price = mean(price),
  mean_price_vol = mean(price_volume),
  mean_units = mean(units)) %>%
  arrange(desc(mean_revenue)) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

* General Mills strongest cereal brand in terms of average revenue is Cheerios.  
  + This is primarily from a high average units combined with a higher price per oz.  
  + Suggesting that Cheerios come in a smaller (lower volume) box or cup, so customers buy more units.  
* In terms of average revenue General Mills Cereals are performing well amoung their peers.

###Ad
```{r}
mtp %>% 
  group_by(company, ad) %>% 
  summarise(count = n(),
  mean_revenue = mean(revenue),
  mean_price = mean(price),
  mean_price_vol = mean(price_volume),
  mean_units = mean(units)) %>%
  arrange(desc(mean_revenue)) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```
* In terms of average revenue General Mills does better than their peers whether implimenting an Ad or not.  
  + Interestingly, no ad performs better than having just a medium/small ad.  
  + The medium/small ads for General Mills are underperforming compared to their peers.  
  + This could be due to the fact that the average price and average price per oz are considerably lower with a B ad.


###Promo
```{r}
mtp %>% 
  group_by(company, promo) %>% 
  summarise(count = n(),
  mean_revenue = mean(revenue),
  mean_price = mean(price),
  mean_price_vol = mean(price_volume),
  mean_units = mean(units)) %>%
  arrange(desc(mean_revenue)) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```
* In terms of average revenue General Mills is performing well compared to their peers whether they are running an instore promo or not.

###Ad and Promo
```{r}
mtp %>% 
  group_by(company, ad, promo) %>% 
  summarise(count = n(),
  mean_revenue = mean(revenue),
  mean_price = mean(price),
  mean_price_vol = mean(price_volume),
  mean_units = mean(units)) %>%
  arrange(desc(mean_revenue)) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

* We know that General Mills' medium/small ads are underperforming.  
  + The underperfoming ads appear to be only when combined with instore promos.  
  + For Kellogs and Post B ads with instore promos are thier best perfoming combination, yet not for General Mills.  
  + **Why are Medium/Small Ads not woking in combination with instore promotions?**


# Univariate graphical: categorical

##Company
```{r}

grid.arrange(
  #Graph Companies by number of sales.
   mtp %>% 
    group_by(company) %>% 
    summarise(count = n()) %>% 
    ggplot(aes(x = reorder(company, count),
             y = count)) + 
    geom_bar(stat = "identity", fill = "steelblue4",
             width = 0.8) +
    theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
    ggtitle("Number of Sales by Company",
          subtitle = "Kelloggs has the highest number of sales.") +
    theme_classic() +
  theme(axis.text = element_text(face = "bold", size = 10),
        axis.title = element_blank(),
        axis.ticks.x = element_blank(),
        legend.title = element_text(face = "bold")) ,
  
  #Graph Companies by mean revenue.
  mtp %>% 
    group_by(company) %>% 
    summarise(mean_revenue = mean(revenue), count = n()) %>%
    ggplot(aes(x = reorder(company, count),
             y = mean_revenue)) + 
    geom_bar(stat = "identity", fill = "steelblue4",
             width = 0.8) +
    theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
    ggtitle("Average Revenue by Company",
          subtitle = "General Mills has the highest average revenue.") +
    theme_classic() +
    theme(axis.text = element_text(face = "bold", size = 10),
        axis.title = element_blank(),
        axis.ticks.x = element_blank(),
        legend.title = element_text(face = "bold")) ,

ncol = 2
)
```

* We see that Kelloggs leads the idustry in number of sales but General Mills has a higher average revenue.  
  + This is good for General Mills, indicating that there is possible room to grow.  
  + Post lags in both sales and mean revenue.

##Cereal
```{r}
#Graphing the most popular cereal brands by number of sales.
mtp %>% 
    group_by(cereal) %>% 
    summarise(count = n()) %>% 
    ggplot(aes(x = reorder(cereal, count),
             y = count)) + 
    geom_bar(stat = "identity", fill = "steelblue4") +
    theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
    coord_flip() +
    xlab("Cereal") +
    ylab("") +
    ggtitle("Number of Sales by Cereal",
          subtitle = "Frosted Flakes and Froot Loops are best selling cereals") +
  theme_classic() +
  theme(axis.text = element_text(face = "bold", size = 9),
        axis.title.x = element_blank(),
        axis.title = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.title = element_text(face = "bold")) 
```

* Bassed on number of sales, Frosted Flakes and Froot Loops (both Kelloggs cereals) are the most popular cereal brands.  
* For the Genereal Mills brands Cinnamon Toast Crunch is the most popular.


```{r}
#Graphing mean revenue by cereal brand.
mtp %>% 
    group_by(cereal) %>% 
    summarise(mean_revenue = mean(revenue)) %>% 
    ggplot(aes(x = reorder(cereal, mean_revenue),
             y = mean_revenue)) + 
    geom_bar(stat = "identity", fill = "steelblue4") +
    theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
    coord_flip() +
    xlab("Cereal") +
    ylab("") +
    ggtitle("Average Revenue by Cereal",
          subtitle = "Cheerios has a very high average revenue compared to others") +
  theme_classic() +
  theme(axis.text = element_text(face = "bold", size = 9),
        axis.title.x = element_blank(),
        axis.title = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.title = element_text(face = "bold")) 
```
```{r}
#Graphing mean units by cereal
mtp %>% 
    group_by(cereal) %>% 
    summarise(mean_units = mean(units)) %>% 
    ggplot(aes(x = reorder(cereal, mean_units),
             y = mean_units)) + 
    geom_bar(stat = "identity", fill = "steelblue4") +
    theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
    coord_flip() +
    xlab("Cereal") +
    ylab("") +
    ggtitle("Average Units per sale by Cereal",
          subtitle = "Cheerios has a very high average revenue compared to others") +
  theme_classic() +
  theme(axis.text = element_text(face = "bold", size = 9),
        axis.title.x = element_blank(),
        axis.title = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.title = element_text(face = "bold")) 
```

* Graphically we can see what we intuitively know; that high mean units are closely related to high mean revenue.  
  + This relationship makes sense because units is part of the revenue calculation.  
  + Does higher mean units or higher mean price have the most significant affect on mean revenue?
* Cheerios has the highest average revenue per sale and highest units per sale. (General Mills brand).  
  + This likely accounts for much of General Mills success as a company at generating substantial revenues.


##Ad
```{r}
#Graphing mean units and mean revenue by ad, examining ad effectiveness.
grid.arrange(
  mtp %>% 
    mutate(ad = if_else(ad == "A", "Big", if_else(ad == "B", "Medium/Small", "None"))) %>% 
    group_by(ad) %>% 
    summarise(mean_units = mean(units)) %>% 
    ggplot(aes(x = reorder(ad, mean_units),
             y = mean_units)) + 
    geom_bar(stat = "identity", fill = "steelblue4") +
    theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
    ggtitle("Average # of Units Sold by Ad Type",
          subtitle = "What type of Ad moves the most units?") +
    theme_classic() +
    theme(axis.text = element_text(face = "bold", size = 10),
        axis.title = element_blank(),
        axis.ticks.x = element_blank(),
        legend.title = element_text(face = "bold")) ,

  mtp %>% 
    mutate(ad = if_else(ad == "A", "Big", if_else(ad == "B", "Medium/Small", "None"))) %>% 
    group_by(ad) %>% 
    summarise(mean_revenue = mean(revenue)) %>%
    ggplot(aes(x = reorder(ad, mean_revenue),
             y = mean_revenue)) + 
    geom_bar(stat = "identity", fill = "steelblue4") +
    theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
    ggtitle("Average Revenue by Ad Type",
          subtitle = "Do ads lead to larger revenue on average?") +
    theme_classic() +
    theme(axis.text = element_text(face = "bold", size = 10),
        axis.title = element_blank(),
        axis.ticks.x = element_blank(),
        legend.title = element_text(face = "bold")),
  

ncol = 2
)
```

* Running adds looks to slightly increase the average units sold but have very little affect on average revenue.  
* Why are ads not have more effect on mean revenue, while also increasing mean units sold.

##Promo
```{r}
#Graphing mean units and mean revenue for in-store promotions.
grid.arrange(
  mtp %>% 
    mutate(promo = if_else(promo == 1, "Yes", "No")) %>% 
    group_by(promo) %>% 
    summarise(mean_units = mean(units)) %>% 
    ggplot(aes(x = reorder(promo, mean_units),
             y = mean_units)) + 
    geom_bar(stat = "identity", fill = "steelblue4") +
    theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
    xlab("Instore Promos") +
    ylab("") +
    ggtitle("Average # of Units Sold by Promo",
          subtitle = "Running promos increases average units per sale.") +
    theme_classic() +
    theme(axis.text = element_text(face = "bold", size = 10),
        axis.title = element_blank(),
        axis.ticks.x = element_blank(),
        legend.title = element_text(face = "bold")),

  mtp %>% 
    mutate(promo = if_else(promo == 1, "Yes", "No")) %>% 
    group_by(promo) %>% 
    summarise(mean_revenue = mean(revenue)) %>%
    ggplot(aes(x = reorder(promo, mean_revenue),
             y = mean_revenue)) + 
    geom_bar(stat = "identity", fill = "steelblue4") +
    theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
    xlab("Instore Promos") +
    ylab("") +
    ggtitle("Average Revenue by Promo",
          subtitle = "Promos don't lead to much higher average revenue.") +
    theme_classic() +
    theme(axis.text = element_text(face = "bold", size = 10),
        axis.title = element_blank(),
        axis.ticks.x = element_blank(),
        legend.title = element_text(face = "bold")),
  

ncol = 2
)
```

* Similar as with the ads, instore promos look to slightly increase average units but have very little affect on average revenue.  


#Multi-variate graphical: quantitative
## Correlations

```{r, message = FALSE, error = FALSE}
ggpairs(subset(mtp, select = -c(iri_key, package, cereal, company, week)))
```

**Observations on correlations and multi-variate relationships.**  
- As would be expected the correlation between units and revenue is very high, very important factor for increasing revenue.
- There is a negative correlation between price and revenue, so raising prices will decrease revenue.  
- Price looks to be follow a somewhat normal distribution.  
- Units and Revenue appear to have a Chi-squarted distribution.  
- Both ad and promo have a lot of revenue values above the interquartile range, it should be okay but be cautious using mean.  

#Multi-variate graphical: categorical and quantitative

```{r}
mtp %>% 
  mutate(ad = if_else(ad == "A", "Big", if_else(ad == "B", "Medium/Small", "None"))) %>% 
  group_by(cereal, ad) %>%
  summarise(units_sold = sum(units) / 1000) %>%
  ggplot(aes(x = reorder(cereal, units_sold), y = units_sold, fill = ad)) +
    geom_bar(stat = "identity") + 
    labs(fill = "Ad", y = "Units sold (,000)") +
    coord_flip() +
    guides(fill = guide_legend(reverse = TRUE)) +
    theme_classic() +
    theme(legend.position = "top", 
          legend.justification = "center",
          legend.title = element_text(face = "bold"),
          axis.title.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.line.y = element_blank()) +
    ggtitle("Units Sold Per Cereal",
          subtitle = "Most units are not sold in conjuction with ads")
```

```{r}
mtp %>% 
  group_by(cereal, promo) %>% 
  summarise(subtotal = sum(units)) %>% 
  group_by(cereal) %>% 
  mutate(promo_pct = subtotal / sum(subtotal),
         promo = if_else(promo == 1, "Yes", "No")) %>%
  arrange(promo, promo_pct) %>% 
  mutate(brand2 = factor(cereal, levels = filter(., promo == "Yes") %>% .$cereal)) %>% 
  ggplot(aes(x = brand2, y = promo_pct, fill = promo)) +
    geom_bar(stat = "identity") + 
    labs(fill = "Promotion", y = "Sales") +
    coord_flip() +
    guides(fill = guide_legend(reverse = TRUE)) +
    scale_y_continuous(labels = percent_format(accuracy = 1, suffix = "%")) +
    theme_classic() +
    theme(legend.position = "top", 
          legend.justification = "center",
          legend.title = element_text(face = "bold"),
          axis.title.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.line.y = element_blank()) +
    ggtitle("Promotional Distribution",
          subtitle = "Cheerios are not run on promotion very often")
```

#Research Questions

##Question 1: How effective is Genereal Mills' promotional strategy and what brands can we focus on to improve promotional success? 

###First we examine General Mills' current promotional strategy compared to the market.

```{r}
mtp %>% 
  filter(promo == 1) %>%
  group_by(company) %>% 
  summarise(Count = n(),
  Average_Revenue = mean(revenue),
  Average_Price = mean(price),
  Average_Price_Vol = mean(price_volume),
  Average_Units = mean(units)) %>%
  arrange(desc(Average_Revenue)) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  row_spec(1, background = "#e5f5e0")
```

**Null Hypothesis: Revenue is the same across all companies when running instore promotions.**

```{r}
mtp_hyp1 <- mtp %>%
  filter(promo == 1) 

# Is revenue the same across brands?
t.test(mtp_hyp1$revenue[mtp_hyp1$company == "General_Mills"], mtp_hyp1$revenue[mtp_hyp1$company == "Kelloggs"])

t.test(mtp_hyp1$revenue[mtp_hyp1$company == "General_Mills"], mtp_hyp1$revenue[mtp_hyp1$company == "Post"])

```

```{r}
#Graphing the relationship between in-store promotions and revenue across companies.
# 90% CI, get z-value for upper tail, use .95 since is one sided
z <- qnorm(.95)

# Incorporate CI into bar graph of means
mtp_hyp1 %>%
  group_by(company) %>%
  summarise(m = mean(revenue), sd = sd(revenue), 
            n = n(), ci = z * sd/sqrt(n)) %>%
  ggplot(aes(x = company, y = m)) +
  geom_bar(stat = "identity", position = "dodge", fill = "#9ecae1") +
  geom_errorbar(aes(ymin = m - ci, ymax = m + ci), 
                width = 0.5, position = position_dodge(0.9)) +
  theme_classic() +
  xlab("") +
  ylab("") +
  theme(axis.ticks.x = element_blank(),
          axis.title.x = element_blank(),
          axis.ticks.y = element_blank()) +
  ggtitle("Mean Revenue When Running a Promotion",
          subtitle = "General Mills' promotions have higher average than competition.")
```

**Findings**  
* Based on our sample and the above analysis we can see that General Mills current promotional strategy fairs better than the rest of the market.  
* Examining the t-tests we can reliably say that in terms of mean revenue General Mills' results are statistically different.  
* Current strategy seems to be working well in comparison to peers.  

###Next we examine General Mills' family of cereal brands to see if we can improve promotional performance anywhere.

```{r}
mtp_gm <- mtp %>%
  filter(company == "General_Mills")

#Create a table of General Mills cereals and their mean revenue when an in-store promotion is happening.
mtp_t1 <- mtp_gm %>% 
  filter(promo == 1) %>%
  group_by(cereal) %>% 
  summarise(Count = n(),
  Average_Revenue = mean(revenue),
  Average_Price = mean(price),
  Average_Price_Vol = mean(price_volume),
  Average_Units = mean(units)) %>%
  arrange(desc(Average_Revenue)) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  row_spec(5, background = "#fee0d2")

mtp_t1

```
Analysis

* When on promotion Cheerios does very well, leading the industry in mean revenue and mean units.  
* Both Cocoa Puffs and Kix are struggling with mean revenue both during promotion and not.  
  + This is likely due to lot mean units, perhaps the price is to high.  
  + Of interest, the average price per volume is very high for Kix, too small of package?  
  

**Null Hypothesis: Revenue is the same across all General Mills cereals during in-store promotions.**

```{r}
mtp_hyp2 <- mtp_gm %>%
  filter(promo == 1)

# Is mean revenue the same across all General Mills' brands?

t.test(mtp_hyp2$revenue[mtp_hyp2$cereal == "Kix"], mtp_hyp2$revenue[mtp_hyp2$cereal == "Cheerios"])

t.test(mtp_hyp2$revenue[mtp_hyp2$cereal == "Kix"], mtp_hyp2$revenue[mtp_hyp2$cereal == "Cinnamon Toast Crunch"])

t.test(mtp_hyp2$revenue[mtp_hyp2$cereal == "Kix"], mtp_hyp2$revenue[mtp_hyp2$cereal == "Cocoa Puffs"])

t.test(mtp_hyp2$revenue[mtp_hyp2$cereal == "Kix"], mtp_hyp2$revenue[mtp_hyp2$cereal == "Lucky Charms"])
```
```{r}
#Graphing mean promotional revenue across General Mill branded cereals.
# 90% CI, get z-value for upper tail, use .95 since is one sided
z <- qnorm(.95)

# Incorporate CI into bar graph of means
mtp_gm %>%
  mutate(promo = if_else(promo == "0", "None", "In-store Promotion")) %>%
  group_by(cereal, promo) %>%
  summarise(m = mean(revenue), sd = sd(revenue), 
            n = n(), ci = z * sd/sqrt(n)) %>%
  ggplot(aes(x = reorder(cereal, m), y = m, fill = promo)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = m - ci, ymax = m + ci), 
                width = 0.5, position = position_dodge(0.9)) +
  coord_flip() +
  theme_classic() +
  xlab("") +
  labs(fill = "Promotion") +
  theme(axis.text = element_text(face = "bold", size = 11),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title = element_blank(),
        legend.position = "top",
        legend.title = element_text(face = "bold")) +
  scale_fill_brewer() +
  ggtitle("Mean Revenue Based on Promotions",
          subtitle = "Cocoa Puffs and Kix lag behind other brands")

mtp_g2 <- mtp_gm %>%
  mutate(promo = if_else(promo == "0", "None", "In-store Promotion")) %>%
  group_by(cereal, promo) %>%
  summarise(m = mean(revenue), sd = sd(revenue), 
            n = n(), ci = z * sd/sqrt(n)) %>%
  ggplot(aes(x = reorder(cereal, m), y = m, fill = promo)) +
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip() +
  scale_y_continuous(limits = c(0, 60),
                     labels = dollar_format(prefix = "$")) +
  theme_classic() +
  labs(fill = "Promotion") +
  theme(axis.text = element_text(face = "bold", size = 11),
        axis.title = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line.y = element_blank(),
        axis.line.x = element_line(colour = "grey"),
        legend.position = "top",
        legend.title = element_text(face = "bold")) +
  scale_fill_brewer() +
  ggtitle("Average Revenue Based on Promotions",
          subtitle = "Cocoa Puffs and Kix lag behind other brands")

mtp_g2

ggsave("mtp_g2.png", plot = mtp_g2)

```


```{r}
gPost <- mtp_hyp1 %>%
    filter(company == "Post") %>%
    group_by(company, cereal) %>%
    summarise(m = mean(revenue), sd = sd(revenue), 
            n = n(), ci = z * sd/sqrt(n)) %>%
    ggplot(aes(x = reorder(cereal, m), y = m)) +
    geom_bar(stat = "identity", position = "dodge", 
             fill = "#deebf7") + 
    coord_flip() +
  scale_y_continuous(limits = c(0, 50)) + 
    theme_classic() +
    theme(axis.line.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.line.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.title = element_blank(),
          axis.text.y = element_text(size = 9),
          axis.text.x = element_blank()) +
    ggtitle("Post") +
  theme(plot.title = element_text(hjust = -0.08, 
                                  vjust = -0.30,
                                  colour = "black",
                                  face = "bold",
                                  size = 11),
        plot.margin = unit(c(0.3,0.3,0.3,0.3), "cm"))

gKelloggs <- mtp_hyp1 %>%
    filter(company == "Kelloggs") %>%
    group_by(company, cereal) %>%
    summarise(m = mean(revenue), sd = sd(revenue), 
            n = n(), ci = z * sd/sqrt(n)) %>%
    ggplot(aes(x = reorder(cereal, m), y = m)) +
    geom_bar(stat = "identity", position = "dodge", 
             fill = "#9ecae1") + 
    coord_flip() +
  scale_y_continuous(limits = c(0, 50)) +
    theme_classic() +
    theme(axis.line.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.line.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.title = element_blank(),
          axis.text.y = element_text(size = 9),
          axis.text.x = element_blank()) +
  ggtitle("Kelloggs") +
  theme(plot.title = element_text(hjust = -0.16, 
                                  vjust = -0.16,
                                  colour = "black",
                                  face = "bold",
                                  size = 11),
        plot.margin = unit(c(0.2,0.2,0.2,0.2), "cm"))

gGMills <- mtp_hyp1 %>%
    filter(company == "General_Mills") %>%
    group_by(company, cereal) %>%
    summarise(m = mean(revenue), sd = sd(revenue), 
            n = n(), ci = z * sd/sqrt(n)) %>%
    ggplot(aes(x = reorder(cereal, m), y = m)) +
    geom_bar(stat = "identity", position = "dodge", fill = "#3182bd") +
    coord_flip() +
  scale_y_continuous(limits = c(0, 50),
                     labels = dollar_format(prefix = "$")) +
    theme_classic() +
    theme(axis.line.y = element_blank(),
          axis.title = element_blank(),
          axis.ticks.y = element_blank(),
          axis.line.x = element_line(colour = "grey"),
          axis.text = element_text(size = 9),
          legend.position = "bottom") +
  ggtitle("General Mills") +
  theme(plot.title = element_text(hjust = -0.24, 
                                  vjust = -0.24,
                                  colour = "black",
                                  face = "bold",
                                  size = 11),
        plot.margin = unit(c(0.2,0.2,0.2,0.2), "cm"))

mtp_g1 <- cowplot::plot_grid(gPost, gKelloggs, gGMills, 
                   align = "v", nrow = 3, 
                   rel_heights = c(0.23, 0.5, 0.4),
                   labels = "Average Revenue During Promotions",
                   label_size = 12,
                   hjust = -0.9,
                   vjust = 1.7)
mtp_g1

ggsave("mtp_g1.png", plot = mtp_g1)
```

###Conclusions to Question 1

Based on our analysis we can reliably say that General Mills' current promotional strategy has the competitive advange in the market when looking at mean revenue. Although this finding is both economically and statistically significant there are also areas in which General Mills can refine and improve their promotional strategy. Upon further analysis we can see that the Cocoa Puffs and Kix brands are not performing as well as other brands in the General Mills family of brands. Cocoa Puffs and Kix both have significantly lower mean revenue and mean units per sale than any of the other brands. Why is this? Pricing may need to be changed, Kix has a very high average price per volume so their packaging may be too small and customers are not seeing it as a "good value" whether it is on promotion or not. Testing pricing or packaging with Kix during an in-store promotion would be a valuable opportunity to acquire sample data sets for future analysis. For Cocoa Puffs, promotions significantly improve thier numbers so increasing the number of promotions invovling Cocoa Puffs could yield better numbers for the brand moving forward.

##Question 2: How effective is General Mills' advertising strategy and how can we improve its revenue generating ability?

###First we can examine if ads have any measurable affect on revenue or units sold.

```{r}
#Table for ad effectiveness across the whole market.
mtp %>% 
  group_by(ad) %>% 
  summarise(Count = n(),
  Average_Revenue = mean(revenue),
  Average_Price = mean(price),
  Average_Price_Vol = mean(price_volume),
  Average_Units = mean(units)) %>%
  arrange(desc(Average_Revenue)) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) 

#Table for just General Mills
mtp_gm %>% 
  group_by(ad) %>% 
  summarise(Count = n(),
  Average_Revenue = mean(revenue),
  Average_Price = mean(price),
  Average_Price_Vol = mean(price_volume),
  Average_Units = mean(units)) %>%
  arrange(desc(Average_Revenue)) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) 

```

**Null Hypothesis: Revenue is the same whether they are running ads or not.**

```{r}

# Is revenue the same whether running an add or not?
t.test(mtp$revenue[mtp$ad == "NONE"], mtp$revenue[mtp$ad != "NONE"])

# 
t.test(mtp_gm$revenue[mtp_gm$ad == "NONE"], mtp_gm$revenue[mtp_gm$ad != "NONE"])
```

```{r}
# 90% CI, get z-value for upper tail, use .95 since is one sided
z <- qnorm(.95)

# Incorporate CI into bar graph of means
mtp %>%
  mutate(ad = if_else(ad == "A", "Big", if_else(ad == "B", "Medium/Small", "None"))) %>% 
  group_by(company, ad) %>%
  summarise(m = mean(revenue), sd = sd(revenue), 
            n = n(), ci = z * sd/sqrt(n)) %>%
  ggplot(aes(x = company, y = m, fill = ad)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = m - ci, ymax = m + ci), 
                width = 0.5, position = position_dodge(0.9)) +
  theme_classic() +
  labs(fill = "Ad") +
  theme(axis.text = element_text(face = "bold", size = 11),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title = element_blank(),
        legend.position = "top",
        legend.title = element_text(face = "bold")) +
  scale_fill_brewer(palette = "Purples") +
  ggtitle("Mean Revenue Based on Ads",
          subtitle = "Running ads does not statistically effect revenue. ")
```

**There is no statistical difference in mean revenue when running an ad or not**


###Next we can see if there is a difference in units when running an ad.


**Null Hypothesis: Revenue is the same whether they are running ads or not.**

```{r}
# Are units the same whether running an ads or not?
t.test(mtp$units[mtp$ad == "NONE"], mtp$units[mtp$ad != "NONE"])

# Are units the same whether running ads or not for General Mills?
t.test(mtp_gm$units[mtp_gm$ad == "NONE"], mtp_gm$units[mtp_gm$ad != "NONE"])
```

```{r}
# 90% CI, get z-value for upper tail, use .95 since is one sided
z <- qnorm(.95)

# Incorporate CI into bar graph of means
mtp %>%
  mutate(ad = if_else(ad == "A", "Big", if_else(ad == "B", "Medium/Small", "None"))) %>% 
  group_by(company, ad) %>%
  summarise(m = mean(units), sd = sd(units), 
            n = n(), ci = z * sd/sqrt(n)) %>%
  ggplot(aes(x = company, y = m, fill = ad)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = m - ci, ymax = m + ci), 
                width = 0.5, position = position_dodge(0.9)) +
  theme_classic() +
  labs(fill = "Ad") +
  theme(axis.text = element_text(face = "bold", size = 11),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title = element_blank(),
        legend.position = "top",
        legend.title = element_text(face = "bold")) +
  scale_fill_brewer(palette = "Purples") +
  ggtitle("Mean Units based on Ads",
          subtitle = "Running ads does statistically effect average units. ")
```

```{r}
mtp_gm %>%
  mutate(ad = if_else(ad == "A", "Big", if_else(ad == "B", "Medium/Small", "None")),
         cereal = str_replace(cereal, "Toast Crunch", "Tst Crn")) %>%
  group_by(cereal, ad) %>%
  summarise(m = mean(units), sd = sd(units), 
            n = n(), ci = z * sd/sqrt(n)) %>%
  ggplot(aes(x = reorder(cereal,m), y = m, fill = ad)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = m - ci, ymax = m + ci), 
                width = 0.5, position = position_dodge(0.9)) +
  theme_classic() +
  labs(fill = "Ad") +
  theme(axis.text = element_text(face = "bold", size = 9),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title = element_blank(),
        legend.position = "top",
        legend.title = element_text(face = "bold")) +
  scale_fill_brewer(palette = "Purples") +
  ggtitle("Mean Units Based on Ads",
          subtitle = "Running ads does statistically effect average units. ")

#Graph for Memo finding 2.
mtp_g3 <- mtp_gm %>%
  mutate(ad = if_else(ad == "A", "Big", if_else(ad == "B", "Medium/Small", "None")),
         cereal = str_replace(cereal, "Toast Crunch", "Tst Crn")) %>%
  group_by(cereal, ad) %>%
  summarise(m = mean(units), sd = sd(units), 
            n = n(), ci = z * sd/sqrt(n)) %>%
  ggplot(aes(x = reorder(cereal,m), y = m, fill = ad)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_classic() +
  scale_y_continuous(limits = c(0, 20)) +
  labs(fill = "Ad") +
  theme(axis.text = element_text(face = "bold", size = 9),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title = element_blank(),
        axis.line = element_line(colour = "grey"),
        legend.position = "top",
        legend.title = element_text(face = "bold")) + 
  scale_fill_brewer(palette = "Purples") +
  ggtitle("Average Units Based on Ads",
          subtitle = "Running ads does statistically effect average units. ")

mtp_g3

#Save graph for Memo
ggsave("mtp_g3.png", plot = mtp_g3)
```

###Conclusion to Qustion 2

Based on our analysis of the effectiveness of advertising on Genereal Mills brands we found that there is no statistical difference in effect on revenue when using ads vs. when not using them. This has interesting implications because we can show that advertising does have an effect on units-- increasing them by a statistically significant amount. Since units and revenue are highly positively correlated we would think that these increases in units would lead to an increase in revenue. However, this is not the case. This is probably due to pricing and running ads concurrently with promotions, we recommend trying to increase the amount of ads run without promotions and revist to see if the ad campaigns are successful at increasing revenue.


#Multiple regression with log(earnings) as the dependent variable.

```{r}
mod_all <- lm(log(revenue) ~ cereal + ad + promo + units, data = mtp)

# Review output
summary(mod_all)

mtp_gm <- mtp %>%
  filter(company == "General_Mills")
  
mod_gm <- lm(log(revenue) ~ cereal + ad + promo + units, data = mtp_gm)

# Review output
summary(mod_gm)
```

##Some take-aways from a quick analysis of the regression output.

Note: A regression analysis was not used for our analysis but is included here to give a deeper understanding of the data & dynamics at play, price was excluded to limit endogenous variable bias.  

* We can see the relationship with units; controlling for all other variables revenue increases by roughly 11% for each additional unit and is highly statistically significant.  
  + This is in line with our analysis concerning the importance of units sold on revenue.  

* There is a large negative relationship between running an in-store promotion and revenue; however, we have to remember that this is controlling for all other variables so without an increase in units running a promotion will decrease revenue by 21.65% industry wide and 25.52% for General Mills and is highly statistically significant.  

* Again, emphasizing the importance of units sold, holding all else constant we see that Lucky Charms would increase revenue by roughly 26.2% over Cheerios and is highly statistically significant. This is probably because they are magically delicious. 

* For our research questions these regressions are not extremely helpful but they reinforce a few of the themes present throughout our analysis.
